### 2025-07-23 paper/sections/other/compute.qmd
formatting improved and toc added (to this end, turned appendix into standalone document).

<!-- [@DHPC2022] -->
## Hardware

We performed our experiments on a high-performance cluster. Details about the cluster will be disclosed upon publication to avoid revealing information that might interfere with the double-blind review process. Since our experiments involve highly parallel tasks and rather small models by today's standard, we have relied on distributed computing across multiple central processing units (CPU). Graphical processing units (GPU) were not required.

@@

Model training for the largest grid searches with 270 unique parameter combinations was parallelized across 34 CPUs with 2GB memory each. The time to completion varied by dataset for reasons discussed in @sec-discussion: 0h49m (*Moons*), 1h4m (*Linearly Separable*), 1h49m (*Circles*), 3h52m (*Overlapping*). Model evaluations for large grid searches were parallelized across 20 CPUs with 3GB memory each. Evaluations for all data sets took less than one hour (<1h) to complete.

@@

## Software

All computations were performed in the Julia Programming Language [@bezanson2017julia]. We have developed a package for counterfactual training that leverages and extends the functionality provided by several existing packages, most notably [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) [@altmeyer2023explaining] and the [Flux.jl](https://fluxml.ai/Flux.jl/v0.16/) library for deep learning [@innes2018fashionable;@innes2018flux]. For data-wrangling and presentation-ready tables we relied on [DataFrames.jl](https://dataframes.juliadata.org/v1.7/) [@milan2023dataframes] and [PrettyTables.jl](https://ronisbr.github.io/PrettyTables.jl/v2.4/) [@chagas2024pretty], respectively. For plots and visualizations we used both [Plots.jl](https://docs.juliaplots.org/v1.40/) [@PlotsJL] and [Makie.jl](https://docs.makie.org/v0.22/) [@danisch2021makie], in particular [AlgebraOfGraphics.jl](https://aog.makie.org/v0.9.3/). To distribute computational tasks across multiple processors, we have relied on [MPI.jl](https://juliaparallel.org/MPI.jl/v0.20/) [@byrne2021mpi].

>>>>

<!-- [@DHPC2022] -->
## Hardware {#sec-app-hardware}

We performed our experiments on a high-performance cluster. Details about the cluster will be disclosed upon publication to avoid revealing information that might interfere with the double-blind review process. Since our experiments involve highly parallel tasks and rather small models by today's standard, we have relied on distributed computing across multiple central processing units (CPU). Graphical processing units (GPU) were *not* used.

@@

Model training for the largest grid searches with 270 unique parameter combinations was parallelized across 34 CPUs with 2GB memory each. The time to completion varied by dataset: 0h49m (*Moons*), 1h4m (*Linearly Separable*), 1h49m (*Circles*), 3h52m (*Overlapping*). Model evaluations for large grid searches were parallelized across 20 CPUs with 3GB memory each. Evaluations for all data sets took less than one hour (<1h) to complete but were generally more memory-intensive (see @sec-app-software for additional details)

@@

## Software {#sec-app-software}

All computations were performed in the Julia Programming Language [@bezanson2017julia]. We have developed a package for counterfactual training that leverages and extends the functionality provided by several existing packages, most notably [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) [@altmeyer2023explaining] and the [Flux.jl](https://fluxml.ai/Flux.jl/v0.16/) library for deep learning [@innes2018fashionable;@innes2018flux]. We chose to work with [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) because it currently appears to be the most comprehensive and extensible package for counterfactual explanations. Despite its good interplay with Flux.jl, the package is not, however, optimized to be used in training. This has caused some issues with memory management and bottlenecked performance. All of our code will be provided to reviewers of this work and open-sourced upon publication.

For data-wrangling and presentation-ready tables we relied on [DataFrames.jl](https://dataframes.juliadata.org/v1.7/) [@milan2023dataframes] and [PrettyTables.jl](https://ronisbr.github.io/PrettyTables.jl/v2.4/) [@chagas2024pretty], respectively. For plots and visualizations we used both [Plots.jl](https://docs.juliaplots.org/v1.40/) [@PlotsJL] and [Makie.jl](https://docs.makie.org/v0.22/) [@danisch2021makie], in particular [AlgebraOfGraphics.jl](https://aog.makie.org/v0.9.3/). To distribute computational tasks across multiple processors, we have relied on [MPI.jl](https://juliaparallel.org/MPI.jl/v0.20/) [@byrne2021mpi].

## Reproducibility

We have taken care to set random seeds for reproducibility.
